## Karin/Elizabeth

they have track seg CSV - times are deicmal hours and in LOCAL not UTC

MLD is depth where temp is half degree less than surface
SD is 1 pixel box around
point estimate is just closest

Line 162-164 in segment check my HYCOM expts against their list

For now baja example add 7 (or 8? unclear if dst or now)

They only grab a single time (00:00) each day

Starting to make dataset ready after xmas / early january

Hycom has missing analysis dates? Need to figure out what the error message is
for these so that download doesnt eat shit

http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_19.1?
var=surf_el&var=salinity&var=water_temp&var=water_u&
var=water_v&
north=35.0000&west=229.5&east=254&south=22.5000&
disableProjSubset=on&horizStride=1&
time_start=1996-11-22T00%3A00%3A00Z&time_end=1996-11-29T00%3A00%3A00Z&
timeStride=1&vertCoord=&addLatLon=true&accept=netcdf


## Checking out their data

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(PAMmisc)
source('../../PAMmisc/devel/HYCOMFunctions.R')
segs <- read_csv('BajaStudyArea_Modeling_Segs_forTaiki.csv')
dayFile <- read_csv('BAJA_0.08deg_2018-12-01.csv')

segs <- ekToStandardFormat(segs, offset=7)
oneSeg <- filter(segs, cruiseNum ==unique(segs$cruiseNum[1]))
# oneSeg <- rename(oneSeg, Latitude=mlat, Longitude=mlon)
smallSeg <- head(oneSeg, 10)
```

```{r}
library(PAMmisc)
hy190 <- hycomToEdinfo('GLBu0.08/expt_19.0', chooseVars=FALSE)
hy190 <- varSelect(hy190, c(T, T, T, T,T))
segEnv <- matchEnvData(oneSeg, nc=hy190, var=c('surf_el', 'salinity', 'water_temp', 'water_u', 'water_v'), timeout=360, raw=TRUE, depth=c(0, 250), buffer=c(.08, .08, 0))
```

## Testing ILD calc functions

```{r}
range(smallSeg$UTC)
range(smallSeg$Longitude)
smallSeg$depth <- 0
smallSeg$depth[1] <- 200
ildFile <- 'HY190_ILDTest.nc'
downloadEnv(smallSeg, edinfo=hy190, fileName=ildFile, timeout = 360, buffer=c(.08, .08, 0))
```



```{r}
calcIld <- function(t, depth) {
    if(is.na(t[1])) {
        return(NA)
    }
    t0 <- t[1]
    tml <- t0 - 0.5
    kmax <- length(t[!is.na(t)])
    dmax <- max(depth[!is.na(t)])
    temp1 <- spline(depth[1:kmax], t[1:kmax], n=dmax, xmin=0, xmax=dmax,
                    method='natural')[[2]][1:dmax]
    k <- (which(temp1 < tml))[1]
    max(0, k-1 + (tml-temp1[k-1]) / (temp1[k]-temp1[k-1]))
}

ildMean <- function(x) {
    dVals <- mget('varData', envir=parent.frame(n=3),ifnotfound = NA)$varData$matchDepth
    if(all(is.na(dVals))) {
        return(NA)
    }
    dim(x) <- c(rep(1, 3-length(dim(x))), dim(x))
    ild <- apply(x, c(1, 2), function(t) {
        calcIld(t, dVals)
    })
    if(is.na(ild[2, 2])) {
        return(mean(ild, na.rm=TRUE))
    }
    ild[2, 2]
}

ildSd <- function(x) {
    dVals <- mget('varData', envir=parent.frame(n=3),ifnotfound = NA)$varData$matchDepth
    if(all(is.na(dVals))) {
        return(NA)
    }
    dim(x) <- c(rep(1, 3-length(dim(x))), dim(x))
    ild <- apply(x, c(1, 2), function(t) {
        calcIld(t, dVals)
    })
    sd(ild, na.rm=TRUE)
}

smallSeg <- ncToData(smallSeg, nc=ildFile, var=c('water_temp'), depth=c(0,200), 
                     FUN=list('ildMean'=ildMean, 'ildSd'=ildSd), buffer=c(.08, .08, 0))
ggplot(smallSeg, aes(x=ild.mean, y=water_temp_ildMean)) +
    geom_point() +
    geom_abline(slope=1, intercept=0)
ggplot(smallSeg, aes(x=ild.SD, y=water_temp_ildSd)) +
    geom_point() +
    geom_abline(slope=1, intercept=0)
smallSeg$ild.mean - smallSeg$water_temp_ildMean
```

```{r}
# testing out GLBy 0.04 grid
yseg <- tail(segs, 10)
yseg$UTC <- yseg$UTC + 24*3600 * 30
yseg <- rename(yseg, Latitude=mlat, Longitude=mlon)
yseg$depth <- 0
yseg$depth[1] <- 200
ildyFile <- 'HY93y_ILDTest.nc'
hy93y <- hycomToEdinfo('GLBy0.08/expt_93.0', chooseVars=FALSE)
hy93y <- varSelect(hy93y, c(T,T,T,T,T))
# downloadEnv(yseg, edinfo=hy93y, fileName=ildyFile, timeout = 360, buffer=c(.16, .16, 0))

```

```{r}
rawSeg <- ncToData(yseg, nc=ildyFile, depth=c(0,200), 
                   FUN=list('ildMean'=ildMean, 'ildSd'=ildSd), buffer=c(.16, .16, 0), raw=TRUE)
rawSeg[[1]]$water_temp[,,1]
```
### make 3x3 grid function

```{r}
make33Grid <- function(x) {
    dim(x) <- c(dim(x), rep(1, 3-length(dim(x))))
    dims <- dim(x)
    if(dims[1] == 3 &&
       dims[2] == 3) {
        return(x)
    }
    # 16 buffer returns 5 means .08 grid
    if(dims[1] == 5) {
        x <- x[2:4, , , drop=FALSE]
        dims <- dim(x)
    }
    if(dims[2] == 5) {
        x <- x[, 2:4, , drop=FALSE]
        dims <- dim(x)
    }
    # if a .04 grid, we do averaging of .08 +- .04 *.5
    if(dims[1] == 9) {
        new <- array(NA, dim=c(3, dims[2:3]))
        new[1, , ] <- (.5 * x[2, ,] + .5 * x[4, ,] + x[3, , ]) / 2
        new[2, , ] <- (.5 * x[4, ,] + .5 * x[6, ,] + x[5, , ]) / 2
        new[3, , ] <- (.5 * x[6, ,] + .5 * x[8, ,] + x[7, , ]) / 2
        x <- new
        dims <- dim(x)
    }
    # print(dims)
    if(dims[2] == 9) {
        new <- array(NA, dim=c(dims[1], 3, dims[3]))
        # browser()
        new[, 1, ] <- (.5 * x[, 2, ] + .5 * x[, 4, ] + x[, 3, ]) / 2
        new[, 2, ] <- (.5 * x[, 4, ] + .5 * x[, 6, ] + x[, 5, ]) / 2
        new[, 3, ] <- (.5 * x[, 6, ] + .5 * x[, 8, ] + x[, 7, ]) / 2
        x <- new
    }
    x
}
```
### apply calcs

```{r}
centerMean <- function(x) {
    if(length(dim(x)) == 3) {
        x <- x[, , 1]
    }
    out <- x[2, 2]
    if(is.na(out)) {
        return(mean(x, na.rm=TRUE))
    }
    out
}
calcIld <- function(t, depth) {
    if(is.na(t[1])) {
        return(NA)
    }
    t0 <- t[1]
    tml <- t0 - 0.5
    kmax <- length(t[!is.na(t)])
    dmax <- max(depth[!is.na(t)])
    temp1 <- spline(depth[1:kmax], t[1:kmax], n=dmax, xmin=0, xmax=dmax,
                    method='natural')[[2]][1:dmax]
    k <- (which(temp1 < tml))[1]
    max(0, k-1 + (tml-temp1[k-1]) / (temp1[k]-temp1[k-1]))
}
# to apply to each raw
calcEKParams <- function(x) {
    x$water_temp <- make33Grid(x$water_temp)
    x$water_u <- make33Grid(x$water_u[, , 1])
    x$water_v <- make33Grid(x$water_v[, , 1])
    x$salinity <- make33Grid(x$salinity[, , 1])
    x$surf_el <- make33Grid(x$surf_el)
    x$ild <- apply(x$water_temp, c(1, 2), function(t) {
        calcIld(t, x$matchDepth)
    })
    result <- list(
        sst_mean = centerMean(x$water_temp[, , 1]),
        sst_sd = sd(x$water_temp[, , 1], na.rm=TRUE),
        sss_mean = centerMean(x$salinity),
        sss_sd = sd(x$salinity, na.rm=TRUE),
        u_mean = centerMean(x$water_u),
        u_sd = sd(x$water_u, na.rm=TRUE),
        v_mean = centerMean(x$water_v),
        v_sd = sd(x$water_v, na.rm=TRUE),
        ild_mean = centerMean(x$ild),
        ild_sd = sd(x$ild, na.rm=TRUE),
        ssh_mean = centerMean(x$surf_el),
        ssh_sd = sd(x$surf_el, na.rm=TRUE)
    )
    result
}
```

```{r}
# testing out calc comparison
yseg <- tail(segs, 10)
# yseg$UTC <- yseg$UTC + 24*3600 * 30
yseg <- rename(yseg, Latitude=mlat, Longitude=mlon)
yseg$depth <- 0
yseg$depth[1] <- 200
yseg$UTC <- ceiling_date(yseg$UTC, unit='1day')
ildvFile <- 'HY93v_Test.nc'
# hy93v <- hycomToEdinfo('GLBv0.08/expt_93.0', chooseVars=FALSE)
# hy93v <- varSelect(hy93v, c(T,T,T,T,T))
downloadEnv(yseg, edinfo=hy93v, fileName=ildvFile, timeout = 360, buffer=c(.16, .16, 0))
rawSeg <- ncToData(yseg, nc=ildvFile, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200))
```


```{r}
# calcEKParams(rawSeg[[1]])
# yseg_params <- cbind(yseg, bind_rows(lapply(rawSeg, calcEKParams)))
library(patchwork)
plotEnvComp <- function(x) {
    x$plotId <- 1:nrow(x)
    basicPlot <- ggplot(x, aes(x=plotId)) +
        geom_line(aes(y=sst.mean - sst_mean, col='sst.m')) +
        geom_line(aes(y=sst.SD - sst_sd, col='sst.s')) +
        geom_line(aes(y=sss.mean - sss_mean, col='sss.m')) +
        geom_line(aes(y=sss.SD - sss_sd, col='sss.s')) +
        geom_line(aes(y=u.mean - u_mean, col='u.m')) +
        geom_line(aes(y=u.SD - u_sd, col='u.s')) +
        geom_line(aes(y=v.mean - v_mean, col='v.m')) +
        geom_line(aes(y=v.SD - v_sd, col='v.s')) +
        # geom_line(aes(y=ild.mean - ild_mean, col='ildm')) +
        # geom_line(aes(y=ild.SD - ild_sd, col='ilds')) +
        geom_line(aes(y=ssh.mean - ssh_mean, col='ssh.m')) +
        geom_line(aes(y=ssh.SD - ssh_sd, col='ssh.s')) +
        ggtitle('Basic EnvData Comparison') +
        ylab('Difference (Absolute)')
    ildPlot <- ggplot(x, aes(x=plotId)) +
        geom_line(aes(y=ild.mean - ild_mean, col='ild.m')) +
        geom_line(aes(y=ild.SD - ild_sd, col='ild.s')) +
        ggtitle('ILD Comparison') +
        ylab('Difference (Absolute)')
    # sstmAvg <- median(abs(x$sst.mean), na.rm=TRUE)
    # sstsAvg <- median(abs(x$sst.SD), na.rm=TRUE)
    # sssmAvg <- median(abs(x$sss.mean), na.rm=TRUE)
    # ssssAvg <- median(abs(x$sss.SD), na.rm=TRUE)
    # umAvg <- median(abs(x$u.mean), na.rm=TRUE)
    # usAvg <- median(abs(x$u.SD), na.rm=TRUE)
    # vmAvg <- median(abs(x$v.mean), na.rm=TRUE)
    # vsAvg <- median(abs(x$v.SD), na.rm=TRUE)
    # sshmAvg <- median(abs(x$ssh.mean), na.rm=TRUE)
    # sshsAvg <- median(abs(x$ssh.SD), na.rm=TRUE)
    # ildmAvg <- median(abs(x$ild.mean), na.rm=TRUE)
    # ildsAvg <- median(abs(x$ild.SD), na.rm=TRUE)
    sstmAvg <- diff(range(x$sst.mean, na.rm=TRUE))
    sstsAvg <- diff(range(x$sst.SD, na.rm=TRUE))
    sssmAvg <- diff(range(x$sss.mean, na.rm=TRUE))
    ssssAvg <- diff(range(x$sss.SD, na.rm=TRUE))
    umAvg <- diff(range(x$u.mean, na.rm=TRUE))
    usAvg <- diff(range(x$u.SD, na.rm=TRUE))
    vmAvg <- diff(range(x$v.mean, na.rm=TRUE))
    vsAvg <- diff(range(x$v.SD, na.rm=TRUE))
    sshmAvg <- diff(range(x$ssh.mean, na.rm=TRUE))
    sshsAvg <- diff(range(x$ssh.SD, na.rm=TRUE))
    ildmAvg <- diff(range(x$ild.mean, na.rm=TRUE))
    ildsAvg <- diff(range(x$ild.SD, na.rm=TRUE))
    basicPlotPct <- ggplot(x, aes(x=plotId)) +
        geom_line(aes(y=(sst.mean - sst_mean)/sstmAvg, col='sst.m')) +
        geom_line(aes(y=(sst.SD - sst_sd)/sstsAvg, col='sst.s')) +
        geom_line(aes(y=(sss.mean - sss_mean)/sssmAvg, col='sss.m')) +
        geom_line(aes(y=(sss.SD - sss_sd)/ssssAvg, col='sss.s')) +
        geom_line(aes(y=(u.mean - u_mean)/umAvg, col='u.m')) +
        geom_line(aes(y=(u.SD - u_sd)/usAvg, col='u.s')) +
        geom_line(aes(y=(v.mean - v_mean)/vmAvg, col='v.m')) +
        geom_line(aes(y=(v.SD - v_sd)/vsAvg, col='v.s')) +
        # geom_line(aes(y=ild.mean - ild_mean, col='ildm')) +
        # geom_line(aes(y=ild.SD - ild_sd, col='ilds')) +
        geom_line(aes(y=(ssh.mean - ssh_mean)/sshmAvg, col='ssh.m')) +
        geom_line(aes(y=(ssh.SD - ssh_sd)/sshsAvg, col='ssh.s')) +
        ggtitle('Basic EnvData Comparison') +
        ylab('Difference (Percent)')
    ildPlotPct <- ggplot(x, aes(x=plotId)) +
        geom_line(aes(y=(ild.mean - ild_mean)/ildmAvg, col='ild.m')) +
        geom_line(aes(y=(ild.SD - ild_sd)/ildsAvg, col='ild.s')) +
        ggtitle('ILD Comparison') +
        ylab('Difference (Percent)')
    (basicPlot + ildPlot) / 
        (basicPlotPct + ildPlotPct)
}
# plotEnvComp(yseg_params)
```

## Check how many hycoms

```{r}
hyList <- PAMmisc::hycomList
segs$whichHy <- sapply(segs$UTC, function(t) {
    PAMmisc:::whichHycom(t, hyList)
})
table(segs$whichHy)
```
Do matching test time taken

```{r}
useHy <- 3
seg_one <- segs[segs$whichHy == useHy, ]
seg_one <- rename(seg_one, Latitude = mlat, Longitude = mlon)
seg_one$UTC <- paste0(seg_one$year, '_',
                   seg_one$month, '_',
                   seg_one$day, '_')
seg_one$UTC <- as.POSIXct(seg_one$UTC, format='%Y_%m_%d', tz='UTC') + 24 * 3600
# seg_one <- head(seg_one)
thisHy <- hyList$list[[3]]
thisHy <- varSelect(thisHy, c(T, T, T, T,T))

start <- Sys.time()
raw_one <- matchEnvData(seg_one, nc=thisHy, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=360)
end <- Sys.time()
print(difftime(end, start, units='mins'))
params_one <- bind_rows(lapply(raw_one, calcEKParams))
plotEnvComp(cbind(seg_one, params_one))
```

with GLBu vs GLBv
```{r}
thisHy <- hycomToEdinfo('GLBu0.08/expt_19.1', chooseVars=FALSE)
thisHy <- varSelect(thisHy, c(T, T, T, T,T))

start <- Sys.time()
raw_one_u <- matchEnvData(seg_one, nc=thisHy, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200))
params_one_u <- bind_rows(lapply(raw_one_u, calcEKParams))
end <- Sys.time()
print(difftime(end, start, units='mins'))
plotEnvComp(cbind(seg_one, params_one_u))
```
Adding 2018 GLBv vs GLBy comparison

```{r}
last_seg <- tail(segs, 13)
last_seg <- rename(last_seg, Latitude = mlat, Longitude = mlon)
last_seg$UTC <- ceiling_date(last_seg$UTC, unit='1day') + 24 * 3600
hyV <- hyList$list$`GLBv0.08/expt_93.0`
hyY <- hyList$list$`GLBy0.08/expt_93.0`
hyV <- varSelect(hyV, c(T, T, T, T, T))
hyY <- varSelect(hyY, c(T, T, T, T, T))
yTest <- matchEnvData(last_seg, nc=hyY, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=360)
vTest <- matchEnvData(last_seg, nc=hyV, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=360)
yParam <- calcEKParams(yTest)
vParam <- calcEKParams(vTest)

plotEnvComp(cbind(last_seg, yParam))
plotEnvComp(cbind(last_seg, vParam))
names(vParam) <- c('sst.mean', 'sst.SD', 'sss.mean', 'sss.SD', 'u.mean', 'u.SD', 'v.mean', 'v.SD', 'ild.mean', 'ild.SD',
  'ssh.mean', 'ssh.SD')
plotEnvComp(cbind(last_seg['Idnum'],vParam, yParam))
```
# test y v better

```{r}
one_cruise <- segs[segs$cruiseNum == 1640, ]
one_cruise$UTC <- one_cruise$UTC + 10 * 365 * 24 * 3600
one_cruise <- rename(one_cruise, Latitude = mlat, Longitude = mlon)
hyV <- hyList$list$`GLBv0.08/expt_93.0`
hyY <- hyList$list$`GLBy0.08/expt_93.0`
hyV <- varSelect(hyV, c(T, T, T, T, T))
hyY <- varSelect(hyY, c(T, T, T, T, T))
yTest <- matchEnvData(one_cruise, nc=hyY, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=360)
vTest <- matchEnvData(one_cruise, nc=hyV, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=360)
yParam <- calcEKParams(yTest)
vParam <- calcEKParams(vTest)


names(vParam) <- c('sst.mean', 'sst.SD', 'sss.mean', 'sss.SD', 'u.mean', 'u.SD', 'v.mean', 'v.SD', 'ild.mean', 'ild.SD',
  'ssh.mean', 'ssh.SD')
plotEnvComp(cbind(one_cruise['Idnum'],vParam, yParam))
```

### Test failure scenarios

Ok currently just fails if nc doesnt download.

```{r}
# thisHy <- hycomToEdinfo('GLBu0.08/expt_19.1', chooseVars=FALSE)
# thisHy <- varSelect(thisHy, c(T, T, T, T,T))

start <- Sys.time()
rm(raw_fail)
raw_fail <- matchEnvData(seg_one, nc=thisHy, buffer=c(.16, .16, 0), raw=TRUE, depth=c(0, 200), timeout=2)
# params_one_u <- bind_rows(lapply(raw_one_u, calcEKParams))
end <- Sys.time()
```

## Should we just download files for segs?

Lets try! (nay voice)

```{r}
cruise <- 1509
test_cruise <- segs[segs$cruiseNum == cruise, ]
range(test_cruise$UTC)
length(unique(floor_date(test_cruise$UTC, unit='1day')))
cruise_folder <- paste0(cruise, '_segs')
downloadHYCOM(test_cruise, folder=cruise_folder, offset=0, mode='segment', retry=TRUE, timeout=600)
```
```{r}
testo <- addEnvParams(test_cruise, folder=cruise_folder)
plotEnvComp(testo)
```

Test a fake grid

```{r}
makeLatLongGrid <- function(longRange, latRange, longDiff, latDiff, depth=0, time=nowUTC()) {
    lats <- seq(from=latRange[1], to=latRange[2], by=latDiff)
    lons <- seq(from=longRange[1], to=longRange[2], by=longDiff)
    data.frame(Latitude = rep(lats, length(lons)),
               Longitude = rep(lons, each=length(lats)),
               UTC = time)
}
llg <- makeLatLongGrid(c(-121, -110), c(23, 35), .08, .08, time=min(one_cruise$UTC))
llg$UTC[1] <- min(llg$UTC) + 3 * 24 * 3600
downloadHYCOM(llg, folder='full_test', offset=0, mode='full')

```

```{r}
test_grid <- addEnvParams(llg, folder='full_test')
```

Test on baja full

```{r}
bajaFull <- read_csv('BAJA_0.08deg_2018-12-01.csv')
bajaFull <- rename(bajaFull, Latitude=lat, Longitude=lon180)
bajaFull$UTC <- as.POSIXct('2018-12-01', format='%Y-%m-%d', tz='UTC') + 1 * 24 * 3600 # date refers to local?
noY <- hyList
noY$list <- noY$list[-29]
downloadHYCOM(bajaFull, folder='baja_full', offset=0, mode='full', hyList = noY)

```

```{r}
test_baja <- addEnvParams(bajaFull, folder='baja_full', offset=0)
plotEnvComp(test_baja)
```


